---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svg"
)
```

# hhjboot

<!-- badges: start -->
[![R-CMD-check](https://github.com/Alec-Stashevsky/hhjboot/workflows/R-CMD-check/badge.svg)](https://github.com/Alec-Stashevsky/hhjboot/actions) [![License: GPL v2](https://img.shields.io/badge/License-GPL%20v2-blue.svg)](https://www.gnu.org/licenses/old-licenses/gpl-2.0.en.html)
[![Codecov test coverage](https://codecov.io/gh/Alec-Stashevsky/hhjboot/branch/main/graph/badge.svg)](https://codecov.io/gh/Alec-Stashevsky/hhjboot?branch=main)
<!-- badges: end -->

Regular bootstrap methods rely on assumptions that observations are independent and identically distributed (*i.i.d.*), but this assumption fails for many types of time series because we would expect the observation in the previous period to have some explanatory power over the current observation. This could occur in any time series from unemployment rates, stock prices, biological data, etc. A time series that is *i.i.d.* would look like white noise, since the following observation would be totally independent of the previous one.

To get around this problem, we can retain some of this *time-dependence* by breaking-up a time series into a number of blocks with length *l*. Instead of sampling each observation randomly (with replacement) like a regular bootstrap, we can resample these *blocks* at random. This way within each block the time-dependence is preserved.

The problem with the block bootstrap is the high sensitivity to the choice of block-length, or the number of blocks to break the time series into. 

The goal of `hhjboot` is to simplify and automate the process of selecting a block-length to perform a block bootstrap. `hhjboot` takes its name from the [Hall, Horowitz, and Jing (1995)](https://academic.oup.com/biomet/article-abstract/82/3/561/260651?redirectedFrom=fulltext) method to algorithmically select the optimal block-length for a given stationary time series.

Under the hood, `hhjboot` uses the moving block bootstrap (MBB) procedure according to [Künsch (1989)](https://projecteuclid.org/euclid.aos/1176347265) which resamples blocks from a set of overlapping sub-samples with a fixed block-length. However, the results of `hhjboot` may be generalized to other block bootstrap procedures such as the *stationary bootstrap* of [Politis and Romano (1994).](https://www.tandfonline.com/doi/abs/10.1080/01621459.1994.10476870)


## Installation

<!-- You can install the released version of hhjboot from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("hhjboot") -->
<!-- ``` -->

Install the development version from [GitHub](https://github.com) with:

``` r
# install.packages("devtools")
devtools::install_github("Alec-Stashevsky/hhjboot")
```

## Upcoming changes
- change parallel to `{foreach}` (Pending feedback)
- Build tests
  - Overlapping subsamples cover entire series
  

## Example

We want to select the optimal block-length to perform a block bootstrap on a 
simulated autoregressive *AR(1)* time series.

First we will generate the time series:
```{r series}
library(hhjboot)
set.seed(32)

# Simulate AR(1) time series
series <- stats::arima.sim(model = list(order = c(1, 0, 0), ar = 0.5),
                           n = 500, rand.gen = rnorm)
```

Now, we can find the optimal block-length to perform a block-bootstrap. We do 
this using two methods.

#### 1. The Hall, Horowitz, and Jing (1995) "HHJ" Method

```{r hhj}
## Using the HHJ Algorithm with overlapping subsamples of width 10
hhjboot(series, sub_block_length = 10)
```


#### 2. The Politis and White (2004) Spectral Density Estimation "PWSD" Method

```{r pwsd}
# Using Politis and White (2004) Spectral Density Estimation
pwsd(series) 
```

We can see that both methods produce similar results for a block-length of 10-11
depending on the type of bootstrap method used.
