---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hhjboot

<!-- badges: start -->
[![R-CMD-check](https://github.com/Alec-Stashevsky/hhjboot/workflows/R-CMD-check/badge.svg)](https://github.com/Alec-Stashevsky/hhjboot/actions)[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing) [![GPL license](https://img.shields.io/badge/License-GPL-blue.svg)](http://perso.crans.org/besson/LICENSE.html)
<!-- badges: end -->

The goal of hhjboot is to simplify and automate the process of selecting a block length to perform a moving block bootstrap (MBB). hhjboot takes its name from the Hall, Horowitz, and Jing (1995) method to algorithmically select the optimal block length for the moving block bootstrap on a given time series.

## Installation

<!-- You can install the released version of hhjboot from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("hhjboot") -->
<!-- ``` -->

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Alec-Stashevsky/hhjboot")
```

## Notes
- change parallel package to suggest instead of import and add requireNamespace()
- build test to make sure overlapping subsamples cover entire series

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(hhjboot)
set.seed(32)

# Simulate AR(1) time series
series <- stats::arima.sim(model = list(order = c(1, 0, 0), ar = 0.5),
                           n = 500, rand.gen = rnorm)

# Find optimal block length using overlapping subsamples of width 10
hhjboot(series, sub.block.size = 10)
```
