---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hhjboot

<!-- badges: start -->
<!-- badges: end -->

The goal of hhjboot is to simplify and automate the process of selecting a block length to perform a moving block bootstrap (MBB). hhjboot takes its name from the Hall, Horowitz, and Jing (1995) method to algorithmically select the optimal block length for the moving block bootstrap on a given time series.

## Installation

<!-- You can install the released version of hhjboot from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("hhjboot") -->
<!-- ``` -->

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Alec-Stashevsky/hhjboot")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(hhjboot)

# Simulate AR(1) time series
series <- stats::arima.sim(model = list(order = c(1, 0, 0), ar = 0.5),
                                     n = 500, innov = rnorm(500))
# Default
HHJboot(series)

# With parallel grid-search
cl <- parallel::makeCluster(parallel::detectCores())
HHJboot(series, cl = cl)

# With given pilot block length and size of overlapping subsample
HHJboot(series, pilot.block.length = 15, sub.block.size = 60)

```
